SI: Settlements

### 2021-05-28 analysis/paper/paper.Rmd
finish dataprep, cont. analysis

```{r setup, echo=FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  cache = TRUE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)
```

@@

```{r plot-rm-dist-min, fig.cap="Min. distance to raw material sources."}
rm_dist %>%
  ggplot(aes(x = min, y = rm, fill = region)) +
  geom_violin() +
  scale_fill_brewer(palette = "Accent") +
  facet_wrap(~type, scales = "free_y") +
  labs(x = "min. dist. (km)", y = "raw material", fill = "Region") +
  theme_minimal() +
  theme(legend.position = "bottom")
```
## Settlement density against distance to raw materials

@@

```{r plot-rm-dist-kde-chrono1, fig.cap="KDE against distance to raw material source, Morava river catchment, pottery traditions."}
kde1 %>% select(id, period_label, kde) %>% 
  left_join(rm_dist) %>% 
  mutate(region = str_extract(id, "^.")) %>% 
  filter(rm %in% c("ad", "zelesice", "pmd", "kl"),
         region == "M") %>% 
  ggplot(aes(mean, kde)) +
  geom_point(alpha = 0.1) + 
  stat_smooth(method = "lm", se = FALSE) +
  facet_grid(vars(period_label), vars(rm)) +
  labs(x = "Mean distance to raw material source (km)",
       y = "KDE") +
  theme_minimal() +
  theme(strip.text.y = element_text(angle = 0))
```

@@

```{r plot-rm-dist-kde-chrono2, fig.cap="KDE against distance to raw material source, Morava river catchment, pottery groups."}
kde2 %>% select(id, period_label, kde) %>% 
  left_join(rm_dist) %>% 
  mutate(region = str_extract(id, "^.")) %>% 
  filter(rm %in% c("ad", "zelesice", "pmd", "kl"),
         region == "M") %>% 
  ggplot(aes(mean, kde)) +
  geom_point(alpha = 0.1) + 
  stat_smooth(method = "lm", se = FALSE) +
  facet_grid(vars(period_label), vars(rm)) +
  labs(x = "Mean distance to raw material source (km)",
       y = "KDE") +
  theme_minimal() +
  theme(strip.text.y = element_text(angle = 0))
```
```{r tab-rm-dist-cor-kde2}
kde2 %>% select(id, period_label, kde) %>% 
  left_join(rm_dist) %>% 
  mutate(region = str_extract(id, "^.")) %>% 
  filter(rm %in% c("ad", "zelesice", "pmd", "kl"),
         region == "M") %>% 
  group_by(period_label, rm) %>% 
  summarise(cor = cor(min, kde)) %>% 
  pivot_wider(id_cols = period_label, names_from = rm, values_from = cor) %>% 
  rename(period = period_label) %>% 
  knitr::kable(caption = "Table of correlations between minimal distance to raw material and settlement density, pottery groups.") %>% 
  kableExtra::kable_styling(full_width = FALSE)
```

@@

@@

```{r get-raster}
hydro_dens <- raster::raster(
  here::here(derived_data, "hydro", "density_raster.tif"))

hydro_dens_at_points <- hydro_dens %>% 
  raster::extract(set_spat) %>% 
  as_tibble() %>% 
  rename(hydro_kde = value) %>% 
  bind_cols(select(st_drop_geometry(set_spat), id))

res1_long <- set_base$period1 %>% 
  select(-ends_with("label")) %>% 
  left_join(hydro_dens_at_points) %>% 
  pivot_longer(cols = hydro_kde) %>% 
  rename(variable = name) %>% 
  bind_rows(res1_long)

res2_long <- set_base$period2 %>% 
  select(-ends_with("label"), -facet) %>% 
  left_join(hydro_dens_at_points) %>% 
  pivot_longer(cols = hydro_kde) %>% 
  rename(variable = name) %>% 
  bind_rows(res2_long)
```

```{r plot-hydro-density, fig.cap="Kernel smoothed intensity of watercourses overlaid with sites (raster clipped to research area extent)."}
hydro_dens %>% 
  raster::mask(mask) %>% 
  raster::rasterToPoints() %>%
  as_tibble() %>%
  rename(KDE = density_raster) %>% 
  ggplot() +
  geom_raster(aes(x, y, fill = KDE)) +
  geom_sf(data = set_spat, alpha = 0.1, shape = 1, color = "white") +
  theme_void()
```

@@

We focus on slope around the settlements

@@

@@

```{r results-export}
res1_long %>% 
  pivot_wider(names_from = variable, values_from = value) %>% 
  arrange(id, period, chrono) %>% 
  write_csv(here::here(derived_data, "results", "results1.csv"))

res2_long %>% 
  pivot_wider(names_from = variable, values_from = value) %>% 
  arrange(id, period, chrono) %>% 
  write_csv(here::here(derived_data, "results", "results2.csv"))
```

>>>>

```{r setup, echo=FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  cache = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)
```

@@

<!-- ```{r plot-rm-dist-min, fig.cap="Min. distance to raw material sources."} -->
<!-- rm_dist %>% -->
<!--   ggplot(aes(x = min, y = rm, fill = region)) + -->
<!--   geom_violin() + -->
<!--   scale_fill_brewer(palette = "Accent") + -->
<!--   facet_wrap(~type, scales = "free_y") + -->
<!--   labs(x = "min. dist. (km)", y = "raw material", fill = "Region") + -->
<!--   theme_minimal() + -->
<!--   theme(legend.position = "bottom",  -->
<!--         panel.border = element_rect(color = "black", fill = NA)) -->
<!-- ``` -->

## Settlement density against distance to raw materials

@@

```{r plot-rm-dist-kde-chrono1, fig.cap="KDE against distance to raw material source, Morava river catchment, pottery traditions."}
kde1 %>% select(id, period_label, kde) %>% 
  left_join(rm_dist) %>% 
  mutate(region = str_extract(id, "^.")) %>% 
  filter(rm %in% c("ad", "zelesice", "pmd", "kl"),
         region == "M") %>% 
  ggplot(aes(mean, kde)) +
  geom_point(alpha = 0.1) + 
  stat_smooth(method = "lm", se = FALSE) +
  facet_grid(vars(period_label), vars(rm)) +
  labs(x = "Mean distance to raw material source (km)",
       y = "KDE") +
  theme_minimal() +
  theme(strip.text.y = element_text(angle = 0),
        panel.border = element_rect(color = "black", fill = NA))
```

@@

```{r plot-rm-dist-kde-chrono2, fig.cap="KDE against distance to raw material source, Morava river catchment, pottery groups."}
kde2 %>% select(id, period_label, kde) %>% 
  left_join(rm_dist) %>% 
  mutate(region = str_extract(id, "^.")) %>% 
  filter(rm %in% c("ad", "zelesice", "pmd", "kl"),
         region == "M") %>% 
  ggplot(aes(mean, kde)) +
  geom_point(alpha = 0.1) + 
  stat_smooth(method = "lm", se = FALSE) +
  facet_grid(vars(period_label), vars(rm)) +
  labs(x = "Mean distance to raw material source (km)",
       y = "KDE") +
  theme_minimal() +
  theme(strip.text.y = element_text(angle = 0),
        panel.border = element_rect(color = "black", fill = NA))
```
```{r tab-rm-dist-cor-kde2}
kde2 %>% select(id, period_label, kde) %>% 
  left_join(rm_dist) %>% 
  mutate(region = str_extract(id, "^.")) %>% 
  filter(rm %in% c("ad", "zelesice", "pmd", "kl"),
         region == "M") %>% 
  group_by(period_label, rm) %>% 
  summarise(cor = cor(min, kde)) %>% 
  pivot_wider(id_cols = period_label, names_from = rm, values_from = cor) %>% 
  rename(period = period_label) %>% 
  knitr::kable(caption = "Table of correlations between minimal distance to raw material and settlement density, pottery groups.") %>% 
  kableExtra::kable_styling(full_width = FALSE)
```

@@

## RM relevant to given period and/or region

```{r rm-period}
# 0 = not used
# 0.5 = somewhat used
# 1 = used
rm_b <- readxl::read_xlsx(here::here("analysis/data/raw_data",
                                     "sidliste_vs_zdroje.xlsx"), sheet = "B")
rm_m <- readxl::read_xlsx(here::here("analysis/data/raw_data",
                                     "sidliste_vs_zdroje.xlsx"), sheet = "M")

rm_relevance <- bind_rows(b = rm_b, m = rm_m, .id = "reg") %>% 
  mutate(across(starts_with("p"), function(x) x * 0.5)) %>% 
  pivot_longer(starts_with("p"), names_to = "period", values_to = "relevance") %>% 
  rename(rm = surovina)

rm_dist_rel_p1 <- rm_dist %>% group_by(id, rm) %>% 
  summarise(mean = mean(mean)) %>% 
  mutate(reg = if_else(str_detect(id, "^B"), "b", "m")) %>% 
  left_join(select(set_base$period1, period, id)) %>% 
  left_join(rm_relevance) %>% 
  filter(relevance > 0) %>% 
  mutate(relevance = factor(relevance, levels = c("0", "0.5", "1")),
         label = labs_chrono$periods[period],
         label = factor(label, levels = labs_chrono$periods))

rm_dist_rel_p2 <- rm_dist %>% group_by(id, rm) %>% 
  summarise(mean = mean(mean)) %>% 
  mutate(reg = if_else(str_detect(id, "^B"), "b", "m")) %>% 
  left_join(select(set_base$period2, period, id)) %>% 
  left_join(rm_relevance) %>% 
  filter(relevance > 0) %>% 
  mutate(relevance = factor(relevance, levels = c("0", "0.5", "1")),
         label = labs_chrono$periods[period],
         label = factor(label, levels = labs_chrono$periods))
```

```{r fun-plot-rm-reg-dist}
plot_rm_reg_dist <- function(x, region) {
  x %>% 
    filter(reg == region) %>% 
    ggplot(aes(label, mean, color = relevance)) +
    geom_violin(show.legend = FALSE) +
    geom_boxplot(width = 0.2, show.legend = FALSE) +
    scale_color_manual(values = c("gray", "black")) +
    facet_wrap(vars(rm), scales = "free_y") + 
    labs(x = element_blank(), y = "Mean distance to raw material source (km)") +
    theme_minimal() + 
    theme(axis.text.x = element_text(angle = 80, hjust = 1))
}
```

### RM relations in *general* pottery traditions

#### Morava river catchment

```{r rm-dist-m-p1, fig.cap="Distance to RM sources, Morava river catchment, pottery traditions. RM sources relevant in given period in black, somewhat relevant RM sources in gray."}
rm_dist_rel_p1 %>% 
  plot_rm_reg_dist("m")
```

#### Eastern part of Bohemia

```{r rm-dist-b-p1, fig.cap="Distance to RM sources, Eastern part of Bohemia, pottery traditions. RM sources relevant in given period in black, somewhat relevant RM sources in gray."}
rm_dist_rel_p1 %>% 
  plot_rm_reg_dist("b")
```

### RM relations in *detailed* pottery groups

#### Morava river catchment

```{r rm-dist-m-p2, fig.cap="Distance to RM sources, Morava river catchment, pottery groups. RM sources relevant in given period in black, somewhat relevant RM sources in gray."}
rm_dist_rel_p2 %>% 
  plot_rm_reg_dist("m")
```

#### Eastern part of Bohemia

```{r rm-dist-b-p2, fig.cap="Distance to RM sources, Eastern part of Bohemia, pottery groups. RM sources relevant in given period in black, somewhat relevant RM sources in gray."}
rm_dist_rel_p2 %>% 
  plot_rm_reg_dist("b")
```

@@

```{r get-raster}
# pkg raster
# hydro_dens <- raster::raster(
#   here::here(derived_data, "hydro", "density_raster.tif"))
# 
# hydro_dens_at_points <- hydro_dens %>% 
#   raster::extract(set_spat) %>% 
#   as_tibble() %>% 
#   rename(hydro_kde = value) %>% 
#   bind_cols(select(st_drop_geometry(set_spat), id))

# pkg stars
hydro_dens <- stars::read_stars(
  here::here(derived_data, "hydro", "density_raster.tif")
) %>% 
  st_set_crs(value = st_crs(set_spat))

hydro_dens_at_points <- hydro_dens %>% 
  stars::st_extract(set_spat) %>% 
  st_drop_geometry() %>% 
  rename(hydro_kde = density_raster.tif) %>% 
  bind_cols(select(st_drop_geometry(set_spat), id))

# bind to results
res1_long <- set_base$period1 %>% 
  select(-ends_with("label")) %>% 
  left_join(hydro_dens_at_points, by = "id") %>% 
  pivot_longer(cols = hydro_kde) %>% 
  rename(variable = name) %>% 
  bind_rows(res1_long)

res2_long <- set_base$period2 %>% 
  select(-ends_with("label"), -facet) %>% 
  left_join(hydro_dens_at_points, by = "id") %>% 
  pivot_longer(cols = hydro_kde) %>% 
  rename(variable = name) %>% 
  bind_rows(res2_long)
```

```{r plot-hydro-density, fig.cap="Kernel smoothed intensity of watercourses overlaid with sites (raster clipped to research area extent)."}
# hydro_dens %>% 
#   raster::mask(mask) %>% 
#   raster::rasterToPoints() %>%
#   as_tibble() %>%
#   rename(KDE = density_raster) %>% 
#   ggplot() +
#   geom_raster(aes(x, y, fill = KDE)) +
#   geom_sf(data = set_spat, alpha = 0.1, shape = 1, color = "white") +
#   theme_void()

ggplot() +
  stars::geom_stars(data = st_crop(hydro_dens, mask), sf = TRUE) +
  geom_sf(data = set_spat, alpha = 0.1, shape = 1, color = "white") +
  theme_void() +
  labs(fill = "KDE")
```
<!--- ====================================================================== -->

@@

We focus on slope around the settlements.

```{r dem-data}
dem_data <- read_csv(here::here(derived_data, "dem/vals_dem.csv"))

# bind to results
res1_long <- set_base$period1 %>% 
  select(-ends_with("label")) %>% 
  left_join(dem_data, by = "id") %>% 
  pivot_longer(cols = c(dem, slope)) %>% 
  rename(variable = name) %>% 
  bind_rows(res1_long)

res2_long <- set_base$period2 %>% 
  select(-ends_with("label"), -facet) %>% 
  left_join(dem_data, by = "id") %>% 
  pivot_longer(cols = c(dem, slope)) %>% 
  rename(variable = name) %>% 
  bind_rows(res2_long)
```

```{r altitude-to-res}
altitude <- set_spat %>% st_drop_geometry() %>% 
  as_tibble() %>% 
  select(id, altitude)

# bind to results
res1_long <- set_base$period1 %>% 
  select(-ends_with("label")) %>% 
  left_join(altitude, by = "id") %>% 
  pivot_longer(cols = altitude) %>% 
  rename(variable = name) %>% 
  bind_rows(res1_long)

res2_long <- set_base$period2 %>% 
  select(-ends_with("label"), -facet) %>% 
  left_join(altitude, by = "id") %>% 
  pivot_longer(cols = altitude) %>% 
  rename(variable = name) %>% 
  bind_rows(res2_long)
```

```{r dem-analysis-correlation, fig.cap='Correlation between altitude derived from DEM and altitude reading in the topographic map.'}
set_spat %>% st_drop_geometry() %>% 
  as_tibble() %>% 
  select(id, altitude) %>% 
  left_join(dem_data) %>% 
  ggplot(aes(altitude, dem)) +
  geom_point() +
  theme_minimal()
```

<!--- ====================================================================== -->

# Settlement 'hierarchy'

Minimal distance to fortified settlements in given regions is explored.

```{r dist-fenced-input}
dist_fenced <- read_rds(here::here(derived_data, "fortif/distance_fortif.RDS"))

# results 1
dist_fenced_p1 <- dist_fenced$traditions %>% 
  mutate(variable = "dist_fenced",
         value = min_dist) %>% 
  select(-min_dist, -reg) %>% 
  left_join(select(set_base$period1, id, period, chrono))

res1_long <- res1_long %>% 
  bind_rows(dist_fenced_p1)

# results 2
dist_fenced_p2 <- dist_fenced$groups %>% 
  mutate(variable = "dist_fenced",
         value = min_dist) %>% 
  select(-min_dist, -reg)

res2_long <- res2_long %>% 
  bind_rows(dist_fenced_p2)
```

<!--- ====================================================================== -->

# Settlement 'continuity'

Is the same location settled in the previous period?

```{r continuity-calc}
settlement_continuity <- function(x) {
  x %>% select(-label, -period_label) %>%
    distinct(id, period, chrono) %>%
    mutate(pres = 1L) %>%
    pivot_wider(names_from = period, values_from = pres, values_fill = 0L) %>%
    pivot_longer(starts_with("p"), names_to = "period") %>%
    mutate(period = factor(period, levels = levels(set_base$period1$period)),
           period_label = labs_chrono$periods[period],
           period_label = factor(period_label, labs_chrono$periods)) %>%
    arrange(id, period, chrono) %>%
    group_by(id, chrono) %>%
    mutate(prev = if_else(value == 1, lag(value), 0L),
           prev = if_else(is.na(prev), 0L, prev),
           reg = if_else(str_detect(id, "^B"), "B", "M")) %>%
    ungroup(id, chrono)
}

cont1 <- set_base$period1 %>%
  settlement_continuity()

cont2 <- set_base$period2 %>%
  settlement_continuity()

# update results
cont1_res <- cont1 %>% select(-value, -reg, -period_label) %>% 
  mutate(variable = "cont") %>% 
  rename(value = prev) %>% 
  filter(value != 0)

res1_long <- res1_long %>% bind_rows(cont1_res)

cont2_res <- cont2 %>% select(-value, -reg, -period_label) %>% 
  mutate(variable = "cont") %>% 
  rename(value = prev) %>% 
  filter(value != 0)

res2_long <- res2_long %>% bind_rows(cont2_res)

rm(list = c("cont1_res", "cont2_res"))
```

```{r continuity-plot1, fig.cap="Continuity of settlements, pottery traditions."}
plot_continuity <- function(x) {
  x %>% 
    group_by(period_label, reg) %>%
    summarise(n = sum(value), prev = sum(prev)) %>%
    mutate(perc = (prev / n) * 100) %>%
    ggplot(aes(period_label, perc, fill = reg)) +
    geom_col(show.legend = FALSE) +
    scale_fill_manual(values = mycol) +
    facet_wrap(vars(reg), ncol = 1) +
    labs(x = "Period", y = "Continuity of settlement (%)") +
    theme_minimal()
}

cont1 %>% plot_continuity()
```
```{r continuity-plot2, fig.cap="Continuity of settlements, pottery groups."}
cont2 %>% plot_continuity()
```

<!--- ====================================================================== -->

# Arrangement along natural lines

Arrangement of settlements along natural 'terrain' lines or along waterways is explored.

```{r along_lines}
along_lines <- read_rds(here::here(derived_data, "lines/linear_arrangement.RDS"))

along_lines1 <- along_lines$traditions %>% 
  mutate(chrono = str_remove(chrono, ".$")) %>% 
  group_by(id, chrono, period, variable) %>% 
  summarise(value = max(value)) %>% 
  ungroup(id, chrono, variable) %>% 
  mutate(value = value / n())

along_lines2 <- along_lines$groups %>% 
  distinct(id, chrono, period, variable, value) %>% 
  group_by(period) %>% 
  mutate(value = value / n())

# results 1
res1_long <- res1_long %>% 
  bind_rows(along_lines1)

# results 2
res2_long <- res2_long %>% 
  bind_rows(along_lines2)
```

<!--- ====================================================================== -->

@@

```{r res-cleaning}
na2null <- function(x) {
  x[is.na(x)] <- 0
  x
}

not_na <- function(x) {
  !is.na(x)
}

# tradice
res1_wide <- res1_long %>% 
  group_by(id, period, variable) %>% 
  summarise(value = mean(value), .groups = "drop") %>% 
  pivot_wider(values_from = value, names_from = variable) %>%
  mutate(across(c(cont, starts_with("line")), na2null)) %>%
  filter(across(everything(), not_na))

res1_long <- res1_wide %>% 
  pivot_longer(-c(id, period), names_to = "variable", values_to = "value")

# skupiny
res2_wide <- res2_long %>% 
  group_by(id, period, variable) %>% 
  summarise(value = mean(value), .groups = "drop") %>% 
  pivot_wider(values_from = value, names_from = variable) %>%
  mutate(across(c(cont, starts_with("line")), na2null)) %>%
  filter(across(everything(), not_na))

res2_long <- res2_wide %>% 
  pivot_longer(-c(id, period), names_to = "variable", values_to = "value")

# scale cols
res1_wide <- res1_wide %>% 
  mutate(across(where(is.numeric), scale))

res2_wide <- res2_wide %>% 
  mutate(across(where(is.numeric), scale))
```

## Violin plots

```{r variable-labels}
labs_variables <- c(
  "settlements_kde" = "Settlement density",
  "dist_fenced" = "Min. dist. to fenced set. (km)",
  "line_terrain" = "Along terrain line",
  "line_water" = "Along water line",
  "cont" = "Settlement continuity",
  "hydro_kde" = "Watercourse density",
  "dem" = "Altitude (DEM)", 
  "altitude" = "Altitude (map)",
  "slope" = "Slope (°)",
  "rm_pc1" = "Distance to RM sources 1",
  "rm_pc2" = "Distance to RM sources 2",
  "rm_pc3" = "Distance to RM sources 3"
) %>% 
  map_chr(str_wrap, 12)
```

```{r syn-p1, fig.height=12}
res1_violin <- res1_long %>% 
  filter(variable != "cont") %>% 
  mutate(reg = str_extract(id, "^."),
         period = labs_chrono$periods[period],
         period = fct_relevel(period, labs_chrono$periods),
         variable = labs_variables[variable],
         variable = fct_relevel(variable, labs_variables)) %>% 
  ggplot(aes(reg, value)) +
  geom_violin(fill = "white") +
  geom_boxplot(aes(fill = reg), width = .2) +
  scale_fill_manual(values = mycol, name = "Region") +
  facet_grid(rows = vars(variable), cols = vars(period), scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "bottom", 
        panel.border = element_rect(color = "black", fill = NA), 
        strip.text.y = element_text(angle = 0))

res1_violin
```

```{r syn-p2, fig.height=12}
res2_violin <- res2_long %>% 
  filter(variable != "cont") %>% 
  mutate(reg = str_extract(id, "^."),
         period = labs_chrono$periods[period],
         period = fct_relevel(period, labs_chrono$periods),
         variable = labs_variables[variable],
         variable = fct_relevel(variable, labs_variables)) %>% 
  ggplot(aes(reg, value)) +
  geom_violin(fill = "white") +
  geom_boxplot(aes(fill = reg), width = .2) +
  scale_fill_manual(values = mycol, name = "Region") +
  facet_grid(rows = vars(variable), cols = vars(period), scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "bottom", 
        panel.border = element_rect(color = "black", fill = NA), 
        strip.text.y = element_text(angle = 0))

res2_violin
```

```{r res-violin-out}
ggsave(here::here("analysis/figures", "res1_violin.png"), res1_violin, 
       width = 9, height = 12)
ggsave(here::here("analysis/figures", "res2_violin.png"), res2_violin, 
       width = 9, height = 12)
```

## Line plots

```{r syn-p1-sum, fig.cap="Summaries for different variables across studied periods. Solid line - mean value, wide transparent line - median value.", fig.height=12}
res1_line <- res1_long %>% 
  mutate(reg = str_extract(id, "^."),
         period = labs_chrono$periods[period],
         period = fct_relevel(period, labs_chrono$periods),
         variable = labs_variables[variable],
         variable = fct_relevel(variable, labs_variables)) %>% 
  group_by(period, variable, reg) %>% 
  summarise(mean = mean(value),
            median = median(value)) %>% 
  ggplot(aes(x = period, color = reg, group = reg)) +
  geom_line(aes(y = mean)) +
  geom_line(aes(y = median), alpha = 0.4, size = 2) +
  scale_color_manual(values = mycol, name = "Region") +
  facet_wrap(vars(variable), scales = "free_y", ncol = 2) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1),
        axis.title = element_blank())

res1_line
```

```{r syn-p2-sum, fig.cap="Summaries for different variables across studied periods. Solid line - mean value, wide transparent line - median value.", fig.height=12}
res2_line <- res2_long %>% 
  mutate(reg = str_extract(id, "^."),
         period = labs_chrono$periods[period],
         period = fct_relevel(period, labs_chrono$periods),
         variable = labs_variables[variable],
         variable = fct_relevel(variable, labs_variables)) %>% 
  group_by(period, variable, reg) %>% 
  summarise(mean = mean(value),
            median = median(value), .groups = "drop") %>% 
  ggplot(aes(x = period, color = reg, group = reg)) +
  geom_line(aes(y = mean)) +
  geom_line(aes(y = median), alpha = 0.4, size = 2) +
  scale_color_manual(values = mycol, name = "Region") +
  facet_wrap(vars(variable), scales = "free_y", ncol = 2) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1),
        axis.title = element_blank())

res2_line
```

```{r res-line-out}
ggsave(here::here("analysis/figures", "res1_line.png"), res1_line, 
       width = 7, height = 7)
ggsave(here::here("analysis/figures", "res2_line.png"), res2_line, 
       width = 7, height = 7)
```

## Analysing results

```{r}
pca2 <- res2_wide %>% 
  group_by(period) %>% 
  nest() %>% 
  mutate(mx = map(data, function(x) select(x, -id)),
         mx = map(mx, as.matrix),
         pca = map(mx, prcomp, center = TRUE),
         sdev = map(pca, function(x) tibble(sdev = x$sdev)),
         sdev = map(sdev, function(x) mutate(x, 
                                             var_prop = sdev^2 / sum(sdev^2),
                                             cum_prop = cumsum(var_prop),
                                             pc = row_number())),
         scores = map(pca, function(x) as_tibble(x$x)),
         rotation = map(pca, broom::tidy, matrix = "rotation"),
         rotation = map(rotation, pivot_wider, names_from = "PC", names_prefix = "PC", values_from = "value")) %>%
  select(-mx, -pca)

pca2 %>% 
  unnest(sdev) %>% 
  ggplot(aes(pc, cum_prop)) +
  geom_col() +
  facet_wrap(~period)

p_pca2_x <- pca2 %>% 
  unnest(c(scores, data)) %>% 
  mutate(reg = str_extract(id, "^.")) %>% 
  ggplot(aes(PC1, PC2, color = reg)) +
  geom_point(alpha = 0.1) +
  facet_wrap(~period, ncol = 1) +
  theme_minimal()

p_pca2_a <- pca2 %>% 
  unnest(c(rotation)) %>% 
  ggplot(aes(PC1, PC2)) +
  geom_segment(xend = 0, yend = 0, alpha = 0.4) +
  geom_text(
    aes(label = column),
    hjust = 1, nudge_x = -0.02, 
    color = "#904C2F",
    size = 2.4
  ) +
  facet_wrap(~period, ncol = 1) +
  theme_minimal()

g_out <- gridExtra::grid.arrange(p_pca2_x, p_pca2_a, ncol = 2)

ggsave(here::here("analysis/figures", "pca2.png"), g_out, height = 21)
```

## Nature vs Culture

## Morava river valley vs East part of Bohemia

<!--- ====================================================================== -->

@@

```{r results-export}
res1_long %>% 
  # distinct(id, chrono, period, variable, .keep_all = TRUE) %>% 
  pivot_wider(names_from = variable, values_from = value) %>% 
  arrange(id, period, chrono) %>% 
  write_csv(here::here(derived_data, "results", "results1.csv"))

# res1_long %>% 
#   add_count(id, chrono, period, variable) %>% 
#   filter(n > 1)

res2_long %>% 
  # distinct(id, chrono, period, variable, .keep_all = TRUE) %>% 
  pivot_wider(names_from = variable, values_from = value) %>% 
  arrange(id, period, chrono) %>% 
  write_csv(here::here(derived_data, "results", "results2.csv"))
```

<!--- ====================================================================== -->
